---
title: "Geospatial Data with R Part 1"
author: "Daniel Kai Sheng Boey"
date: "29th February 2019"
output: 
  html_notebook:
    toc: true
---
## Directories
## TROLOLOLOL
```{r}
setwd("C:/Users/danie/OneDrive/Hertie Spring 20/R Geospatial Data")
library(tidyverse)
library(reshape2)
library(tmap)
library(raster)
library(RColorBrewer)
library(classInt)
library(viridisLite)
library(sp)
library(ggmap)
library(sf)
library(rgdal)
```


# Chapter One: Simple Data using SPDF
Using simple SPDF based on GADM.
```{r}
# Getting SPDF of India
India <- getData("GADM", country = "India", level = 1)
str(India, max.level = 2)
summary(India@data)
tmap_mode("plot")
```
```{r}
# Using Quick Thematic Maps
qtm(India)

# Using Thematic Maps
tm_shape(India)+
  tm_borders()+
  tm_fill(col = "HASC_1")+
  tm_legend(outside = TRUE, outside.position = c("right"))+
tmap_options(max.categories = 36) 
```
# Chapter Two: Netherlands Simple Feature Data Sets
## Simple Analysis
Simple summaries/classes to observe the dataset's characteristics.
```{r}
# Here we see municipal level data of Netherlands
data(NLD_muni)
class(NLD_muni)
data(NLD_prov)
summary(NLD_muni,max.level=2)
```
## Quick Thematic Maps/Quick Plotting with Color Palettes
Lets say we want to see the provinces with high percentages of the population >65 years old. 
```{r}
# Quick Thematic Map
qtm(NLD_muni)
print(NLD_muni) #sp dataset

# Color Palette
yellowred<-brewer.pal(n=9,"YlOrRd")

# Plotting
tm_shape(NLD_muni) + 
  tm_fill("pop_65plus", palette = yellowred,title = "Percentage of County >65 years")+
  tm_borders(alpha = 0.7)+
  tm_layout(title.size = 1)+
    tm_legend(outside = TRUE, outside.position = c('bottom','right'))

# New Theme
tm_shape(NLD_muni) + 
  tm_fill("pop_65plus", palette = yellowred,title = "Percentage of County >65 years")+
  tm_style("cobalt")+
  tm_borders(alpha = 0.5)+
  tm_shape(NLD_prov)+
  tm_borders(lwd = 1,"white")

tmap_save(filename = "test65.html")
```
## Faceting
Here, I facet the data based upon the five population categories. 
```{r}
# Trying faceting
facets = c("pop_0_14", "pop_15_24","pop_25_44","pop_45_64","pop_65plus")
nld_map <-tm_shape(NLD_muni)+
  tm_fill(facets)+
  tm_facets(nrow =2)+
#  tm_basemap(server = "Google")+ 
  tm_style("cobalt")+
  tm_borders(alpha = 0.7)
#  tmap_options(limits = c(facets.view = 5))


#Suppressed to save memory
#nld_map
```
To distinguish among percentages across all, we can use fixed breaks (although it's not very useful here) or quantile breaks. 
```{r}
#Fixed breaks
nld_map2 <-tm_shape(NLD_muni)+
  tm_fill(facets, style = "fixed",breaks = c(5,10,15,20,25,30))+
  tm_facets(nrow =2, sync = TRUE)+
#  tm_basemap(server = "Google")+ 
  tm_style("cobalt")+
  tm_borders(alpha = 0.7)+
  tmap_options(limits = c(facets.view = 5))
#nld_map2

# Quantile breaks
nld_map3 <-tm_shape(NLD_muni)+
  tm_fill(facets, style = "quantile")+
  tm_facets(nrow =2, sync = TRUE)+
#  tm_basemap(server = "Google")+ 
  tm_style("cobalt")+
  tm_borders(alpha = 0.7)+
  tmap_options(limits = c(facets.view = 5))

nld_map3
```
## Data Manipulation + Plotting
What if we want to know the biggest population category in a particular province? First, I pull out the relevant variables in the dataframe within the $sf$. Then, I manipulate to find the population category which has the highest proportion within each municipality.
```{r}
df <- data.frame(NLD_muni[,c("code","pop_0_14", "pop_15_24","pop_25_44","pop_45_64","pop_65plus")])
#str(df)
#Remove geometry and finding the category with the highest pop category
df$geometry <- NULL 
df_m<-df %>%
  dplyr::select("code","pop_0_14", "pop_15_24","pop_25_44","pop_45_64","pop_65plus")%>%
  mutate(high_cat = colnames(df)[apply(df[,!names(df)%in%"code"],1,which.max)+1])

#str(df_m)

#Merge back into dataset
NLD_muni$high_cat <- df_m$high_cat

# New Theme
nld_highpop<- tm_shape(NLD_muni,name = "name") + 
  tm_fill("high_cat",title = "Highest Population Cateogry",labels = c("0-14 years old","25-44 years old","45-64 years old"))+
  tmap_style("col_blind")+
  tm_borders(alpha = 0.5)+
  tm_shape(NLD_prov)+
  tm_borders(lwd = 1,"white")
tmap_mode("view")
nld_highpop
tmap_save(filename = "test_high.html")
```
