---
title: "Geospatial Data with R Part 2"
author: "Daniel Kai Sheng Boey"
date: "29th February 2019"
output: 
  html_notebook:
    toc: true
---

## Directories
```{r}
setwd("C:/Users/danie/OneDrive/Hertie Spring 20/R Geospatial Data")
library(tidyverse)
library(reshape2)
library(tmap)
library(raster)
library(RColorBrewer)
library(classInt)
library(viridisLite)
library(sp)
library(ggmap)
library(sf)
library(rgdal)
```


# Chapter 3: Using ggmap and get_map
## NYC EMS Data
Data taken from: https://data.cityofnewyork.us/Public-Safety/Emergency-Response-Incidents/pasr-j7fb. It details the type and address of emergency incident to which OEM responded.

```{r}
library(lubridate)
df<-read.csv("Emergency_Response_Incidents.csv")
df_s_all<-df[complete.cases(df),] # Obtaining only complete cases
df_s<-df_s_all[1:100,]
str(df_s)
```
## Getting Geospatial Map of NYC
Note: GMap from Google Static Maps API; New York City borough shapefile from: https://data.cityofnewyork.us/City-Government/Borough-Boundaries/tqmj-j8zm

```{r}
### Map file of NYCwith preset key from Google API ###
#Note: OSM is not available
map_nyc <- get_map(location = c(lon=df_s$Longitude[100],lat = df_s$Latitude[100]), zoom = 10)

### For just outlines of states
#library(maps)
#mapnyc<-map('county', 'new york')
#plot(mapnyc)
#map_nyc2<-map_data(mapnyc)

#dir()

### NYC Boroughs Shapefile
shp_nyc_bor<- readOGR("geo_export_c588a074-8667-4647-8342-c5b745d7e53f.shp")
#summary(shp_nyc_bor)

### Confirming projection system of shapefile; default GMaps CRS is WGS84 ###
proj4string(shp_nyc_bor)

### Transforming shapefile into points for ggmap ###
pot_nyc_bor <- fortify(shp_nyc_bor)
#str(pot_nyc_bor)

### Plotting ###
#tiff("nyc_emergency.tiff", width = 4, height = 4, units = 'in', res = 300)
map_1<-ggmap(map_nyc,maprange = TRUE)+
  geom_point(data = df_s,aes(x=Longitude,y=Latitude))+
  geom_polygon(aes(x=long,y=lat, group=group, alpha=0.25), data=pot_nyc_bor, fill='white') +
geom_polygon(aes(x=long,y=lat, group=group), data=pot_nyc_bor, color='black', fill=NA)



map_2<-map_1 + stat_density2d(aes(x = Longitude, y = Latitude, fill = ..level.., alpha =..level..), size = 0.2, bins = 50, data = df_s, geom = "polygon")+
  labs(title ="Density of Emergency Responses in NYC")
```
```{r}
map_2

ggsave("nycemergency.tiff",dpi = 300,width = 4, height = 4, units = c("in"))
```
In the map, we can see that the responses are concentrated in the Bronx, downtown Manhattan, downtown Brooklyn and the middle of Queens in a sample of 100. Whilst the plot may not be very useful for now, I will use further edits to improve the map in the future.

## Sf
```{r}
class(map_nyc)

ggmap_rast = function(map){
  map_bbox = attr(map, 'bb') 
  .extent = extent(as.numeric(map_bbox[c(2,4,1,3)]))
  my_map = raster(.extent, nrow= nrow(map), ncol = ncol(map))
  rgb_cols = setNames(as.data.frame(t(col2rgb(map))), c('red','green','blue'))
  red = my_map
  values(red) = rgb_cols[['red']]
  green = my_map
  values(green) = rgb_cols[['green']]
  blue = my_map
  values(blue) = rgb_cols[['blue']]
  stack(red,green,blue)
}

gr1 = ggmap_rast(map = map_nyc)
plotRGB(gr1)


qtm(gr1)

class(gr1)
bb = bbox(gr1)

# Stitching together many bounding boxes
n_col = 21 # width in bb units ################## add your input here 
n_row = 20 # height ################## add your input here

bb_array = array(NA, dim = c(2, 2, n_row, n_col))
bb_array[1,,1,1] = bb[1, ] - ((n_col - 1) / 2) * (bb[1, 2] - bb[1, 1]) # top left x
bb_array[2,,1,1] = bb[1, ] - ((n_row - 1)  / 2) * (bb[2, 2] - bb[2, 1]) # top left y
for(i in 1:n_col){
  for(j in 1:n_row){
    bb_array[1,,j,i] = bb[1, ] - ((n_col - 1) / 2) * (bb[1, 2] - bb[1, 1]) + (i - 1)* (bb[1, 2] - bb[1, 1])
    bb_array[2,,j,i] = bb[2, ] - ((n_row - 1)  / 2) * (bb[2, 2] - bb[2, 1]) + (j - 1)* (bb[2, 2] - bb[2, 1])
  }
}

# Using the bounding boxes to download the images
grm = ggmap_rast(get_map(bb_array[,,1,1], maptype = "satellite"))
plot(grm)

for(i in 1:n_col){
  for(j in 1:n_row){
    gr = ggmap_rast(get_map(bb_array[,,j,i], maptype = "satellite"))
    grm = raster::merge(grm, gr, tolerance = 5)
  }
}
plotRGB(grm)

writeRaster(grm, filename = "nyc_gmap.tiff", format="GTiff", overwrite = TRUE)

```

