---
title: "Geospatial Data with R Part 3 - SF and Raster Objects"
author: "Daniel Kai Sheng Boey"
date: "3rd March 2019"
output: 
  html_notebook:
    toc: true
---

## Directories
```{r}
setwd("C:/Users/danie/OneDrive/Hertie Spring 20/R Geospatial Data")
library(tidyverse)
library(reshape2)
library(tmap)
library(raster)
library(RColorBrewer)
library(classInt)
library(viridisLite)
library(sp)
library(ggmap)
library(sf)
library(rgdal)
```


# Chapter 4: Manipulating SF and Raster Data


```{r}
library(lubridate)
df<-read.csv("Emergency_Response_Incidents.csv")
df_s_all<-df[complete.cases(df),] # Obtaining only complete cases
df_s<-df_s_all[,]
str(df_s)
```
## Getting Geospatial Map of NYC
Note: GMap from Google Static Maps API; New York City borough shapefile from: https://data.cityofnewyork.us/City-Government/Borough-Boundaries/tqmj-j8zm

```{r}
### Map file of NYCwith preset key from Google API ###
#Note: OSM is not available
map_nyc <- get_map(location = c(lon=df_s$Longitude[100],lat = df_s$Latitude[100]), zoom = 10)

### NYC Neighbourhood Data
shp_nyc_NTA<- st_read("geo_export_e52111a7-ba1f-48e5-876d-c912e4640c4b.shp")
#summary(shp_nyc_bor)

### NYC Boroughs Shapefile
shp_nyc_bor<-st_read("geo_export_5912ccc8-6f97-4ae2-bf31-ce6d35ef2fb4.shp")

### Confirming projection system of shapefile; default GMaps CRS is WGS84 ###
st_crs(shp_nyc_bor)
st_crs(shp_nyc_NTA)

### Creating SF file for database
emr_nyc<-st_as_sf(df_s,coords=c("Longitude","Latitude"))

### Checks for consistent CRS
st_crs(emr_nyc)
st_crs(emr_nyc)<-st_crs(shp_nyc_NTA)
st_crs(shp_nyc_NTA)==st_crs(emr_nyc)

### Initial Plot
plot(st_geometry(shp_nyc_NTA))
plot(emr_nyc, add = TRUE, col = "yellow")



```

## Borough Level Data Analysis
The plot below shows the density of emergency response incidents per 1 million sq km. I begin by using the function $st_intersects()$ and then plot based on each borough. Of the five boroughs, Manhattan has the highest incident density rate, which is not surprising. However, It seems like Manhattan's high rate has eclipsed that of the other boroughs. We can change this by specifying the breaks in the data visualized. 
```{r}
### Calculating Statistics for Borough-Level
count   <- st_intersects(shp_nyc_bor,emr_nyc)
#Counting number of occurences
n_bor   <-lengths(count) 
#Calculating Areas
area_bor<-unclass(st_area(shp_nyc_bor))
#Calculating Density
den_bor <-n_bor/area_bor

comb_bor <- shp_nyc_bor %>%
  mutate(n = n_bor, den = den_bor, area = area_bor, den1mil = den_bor*1e6)
head(shp_nyc_NTA)
### Plotting
maps_bor<- mapview(shp_nyc_NTA,zcol ="ntaname", 
          layer.name = "New York Neighborhoods", 
          lwd = 1,
          color = "black",
          alpha.regions = 0, hide = TRUE,legend=FALSE)+
  mapview(comb_bor, zcol = "den1mil", 
          col.regions = sf.colors(10), 
          alpha.regions = 0.8, 
          layer.name = "Incidents per 1 million sq km",
          at = c(2,5,10,3060))
maps_bor


```


## NTA Level Data Mapping
This map is more useful as we can see finer detail abotu the relative number of incidents, controlling for the area of neighborhoods. Below, we observe that SoHo-TriBeCa-Civic Center-Little Italy has the highest number of incidents. Interestingly, we also see that Kew Gardens in Queens has an unusually high emergency incident rate vis-a-vis other neighborhoods in its surroundings and in Queens. 
```{r}
### Finding out density of incidents per neighborhood
count   <- st_intersects(shp_nyc_NTA,emr_nyc)
#Counting number of occurences
n_res   <-lengths(count) 
#Calculating Areas
area_NTA<-unclass(st_area(shp_nyc_NTA))
#Calculating Density
den_NTA <-n_res/area_NTA

comb_NTA <- shp_nyc_NTA %>%
  mutate(n = n_res, den = den_NTA, area = area_NTA, den1mil = den_NTA*1e6)


### Plotting
# Statistical Viewing
ggplot(comb_NTA)+
  geom_histogram(aes(x = n))
ggplot(comb_NTA, aes(x = n, y = area))+
  geom_point()+
  stat_smooth()
# Using Map View
library(mapview)
maps_NTA<- mapview(shp_nyc_bor,zcol ="boro_name", 
          layer.name = "New York Boroughs", 
          lwd = 5,
          color = "black",
          alpha.regions = 0)+
  mapview(comb_NTA, zcol = "den1mil", 
          col.regions = sf.colors(10), 
          alpha.regions = 0.8, 
          layer.name = "Incidents per 1 million sq km",
          at = seq(0,200,20))

maps_NTA
```
## Using Thematic Maps
```{r}
# Using Thematic Maps
yellowred<-brewer.pal(n=5,"YlOrRd")
tmap_mode("view")
tm_NTA<-tm_shape(comb_NTA)+
  tm_borders()+
  tm_fill(col = "n", title = "Number of Incidents", 
          palette = yellowred, style = "quantile",
          alpha = 0.6)+
  tm_shape(shp_nyc_bor)+
  tm_borders(col = "grey60",lwd = 2)+
  tm_basemap
tm_NTA
tmap_save(filename = "tm_NTA.html")
```

